### 2017-04-06 paper.tex
various fixes

@@

The valid encoding and decoding of a RTTY signal depends on the chosen baud
rate (the number of times the logical signal changes level). During tests,
the highest baud rate that allowed clear decoding of radio signals at 10mW
power rating was 200bauds. Since each byte requires eleven bits to be sent
(one start bit, eight data bits and two stop bits), this corresponds to a
\(\frac{200}{11}=18.18\) bytes per second bandwidth.

@@

\item The Reed-Solomon family of algorithms is widespread in its use, notably
in space communications systems – the Voyager probes used Reed-Solomon codes to
encode their images (\cite{Wicker1994}), the Space Data Links frames (devised
by ESA and CCSDS) use a variant of Reed-Solomon to provide FEC
(\cite{EuropeanCooperationforSpaceStandardization2010}) and the SSDV protocol,
used in High-Altitude Balloon missions to transmit low-resolution images,
uses RS(255,223) with 256-byte frames (\cite{UKHAS2016}).

@@

The chosen solution was to not include the start sequence as part of the
error-corrected message. Since proper detection of that sequence is required
for the decoder to start parsing a frame, it would have been impossible to apply
FEC to a frame whose frame start sequence is corrupted.

@@

Figure \ref{fig:frame-fsm} presents the finite state machine that
\acrshort{ahabus} frame decoders must implement to detect frames in a binary
stream, and figure \ref{fig:frame-fmt-3} the final version of the
\acrshort{ahabus} frame format.

@@

\begin{figure}[H]
    \begin{bytefield}{16}
        \bitheader{0,7,8,15} \\
        \begin{leftwordgroup}{Header}
            \bitbox{8}{\texttt{0x5A}} & \bitbox{8}{version} \\
            \bitbox{16}{sequence number}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox{2}{data[]}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{FEC}
            \wordbox{2}{checksum[32]}
        \end{leftwordgroup}
    \end{bytefield}
    \centering
    \caption{\acrshort{ahabus} Frame Format - Version 3}
    \label{fig:frame-fmt-3}
\end{figure}

@@

Encapsulating the transmitted data adds overhead. If \(n\) is the size of a
chunk in bytes, and each frame adds \(m\) bytes of metadata, then only a
percentage \(\frac{n}{n+m}\) of the stream contains actual data. For reasons
explained further, the \acrshort{ahabus} frames were chosen to be 256-byte
long. As shown previously, each frame also includes.

@@

\begin{itemize}
\item Latitude and Longitudes can be both positive (North, East) or Negative
(South, West). At the equator, \({0.0001}^\circ\) of longitude maps to roughly
11 metres\footnotemark , and that distance decreases when moving towards the
poles, which was deemed precise enough for the project's purpose. To avoid
conflicts in floating-point mathematics conventions, and allow packet encoders
to be written for embedded platform without hardware floating point support,
latitude and longitude were designed to be multiplied by 10,000 and transmitted
as 32-bit precision signed integers.

\item Altitude is above zero for the whole mission duration, and cannot exceed
50,000m. Since a sub-metre precision is not required, and would hardly be
achievable with civilian GNSS receivers, it is encoded as a 16-bit precision
unsigned integer.
\end{itemize}

\footnotetext{
Earth circumference at the equator: \(40\time10^{6}\)m.
\(D_{0.0001^{\circ}} = 0.0001\frac{40\time10^{6}}{360}=11.1...m\)
}

@@

To evaluate the protocol's viability, a reference implementation of an encoder
and decoder was written. The goal was to

\section{Payload Data Bus}

@@

@@

\section{Ground Segment}

@@

\section{Testing}

>>>>

To avoid possible byte order conflicts between the architectures used on the
flight platform and the ground station, multi-bytes values referred to in the
specification should always be transmitted as big-endian (most significant byte
first).

@@

The valid encoding and decoding of a \acrshort{rtty} signal depends on the
chosen baud rate (the number of times the logical signal changes level). During
tests, the highest baud rate that allowed clear decoding of radio signals at
\SI{10}{mW} power rating was \SI{200}{bps}. Since each byte requires eleven
bits to be sent (one start bit, eight data bits and two stop bits), this
corresponds to a \(\frac{200}{11}=\SI{18.18}{Bps}\) bandwidth.

@@

\item The Reed-Solomon family of algorithms is widespread in its use, notably
in space communications systems – the Voyager probes used Reed-Solomon codes to
encode their images (\cite{Wicker1994}), the Space Data Links frames (devised
by \acrshort{esa} and \acrshort{ccsds}) use a variant of Reed-Solomon to
provide \acrshort{fec} (\cite{EuropeanCooperationforSpaceStandardization2010})
and the SSDV protocol, used in High-Altitude Balloon missions to transmit
low-resolution images, uses RS(255,223) with 256-byte frames (\cite{UKHAS2016}).

@@

The chosen solution was to not include the start sequence as part of the
error-corrected message. Since proper detection of that sequence is required
for the decoder to start parsing a frame, it would have been impossible to apply
\acrshort{fec} to a frame whose frame start sequence is corrupted.

@@

Figure \ref{fig:frame-fsm} presents the finite state machine that
AHABus frame decoders must implement to detect frames in a binary
stream, and figure \ref{fig:frame-fmt-3} the final version of the
AHABus frame format.

@@

\begin{figure}[H]
    \begin{bytefield}{16}
        \bitheader{0,7,8,15} \\
        \begin{leftwordgroup}{Header}
            \bitbox{8}{\texttt{0x5A}} & \bitbox{8}{version} \\
            \bitbox{16}{sequence number}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox{2}{data[]}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{FEC}
            \wordbox{2}{checksum[32]}
        \end{leftwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format - Version 3}
    \label{fig:frame-fmt-3}
\end{figure}

@@

Encapsulating the transmitted data adds overhead. If \(n\) is the size of a
chunk in bytes, and each frame adds \(m\) bytes of metadata, then only a
percentage \(\frac{n}{n+m}\) of the stream contains actual data. For reasons
explained further, the AHABus frames were chosen to be 256-byte
long. As shown previously, each frame also includes.

@@

\begin{itemize}
\item Latitude and Longitudes can be both positive (North, East) or Negative
(South, West). At the equator, \SI{0.0001}{\degree} of longitude maps to
roughly eleven metres\footnotemark , and that distance decreases when moving
towards the poles, which was deemed precise enough for the project's purpose.
To avoid conflicts in floating-point mathematics conventions, and allow packet
encoders to be written for embedded platform without hardware floating point
support, latitude and longitude were designed to be multiplied by 10,000 and
transmitted as 32-bit precision signed integers.

\item Altitude is above zero for the whole mission duration, and cannot exceed
\SI{50000}{\meter}. Since a sub-metre precision is not required, and would
hardly be achievable with civilian GNSS receivers, it is encoded as a 16-bit
precision unsigned integer.

\footnotetext{
Earth circumference at the equator: \SI{40e6}{\meter}.
\(D_{\SI{0.0001}{\degree}} = 0.0001\times\frac{40\times{10^6}}{360}=\SI{11.1}{\meter}\)
}

@@

To evaluate the protocol's viability, a full demonstration of an AHABus radio
stack was written. Its main goals were to provide a reference implementation of
a compliant encoder and decoder, validate the protocol's design process, and
test the algorithms used in a controlled environment before integrated testing.

The C language (ISO/IEC 9899:1999 standard) was chosen for the reference
implementation because of its widespread use, and its availability on most
micro-controller and embedded platforms, which would be used as AHABus flight
platforms. To render the code as portable and reusable as possible, care was
taken to limit the dependencies to a subset of the C language standard library,
and avoid use of dynamic memory allocation, which is often inefficient or
unavailable on embedded platforms.

To increase portability and ease of use further, the encoder and decoder read
and write data using user-supplied callback. This allows developers to import
the implementation files and use the library without editing, at the expense of
optimisation.
% TODO: Explain?

\subsubsection{Encoder}

Figure \ref{fig:encoder-flow} illustrate the data flow in the encoder.
In an attempt to use as little memory as possible, no intermediate
representation of the packet is created or stored. Instead, the header and
first chunk of the packet's data is written directly to a frame buffer. Once
the frame is fully constructed, its binary representation is written using the
user-provided callback. The buffer is then cleared, and the process is repeated
until all of the packet's payload has been framed.

\begin{figure}[H]
\includegraphics[width=0.6\textwidth]{encoder-flow.pdf}
\centering
\caption{Encoder Data Flow}
\label{fig:encoder-flow}
\end{figure}

For testing purposes, a simple program was built that takes in user input,
uses the encoder to package that data and outputs the frame stream to the
UNIX standard output (\texttt{stdout}).

% TODO: more details? how much is required?
% TODO: add reference to appendix with code

\subsubsection{Decoder}

\subsubsection{Initial Testing}

\section{Payload Bus}
\label{sec:payload-bus}

The Payload Bus is the hardware and software interface that is used by the
\acrfull{obc} to communicate with each payload carried on a mission.

@@

\subsection{Initial Prototypes}

Before a protocol was chosen, two different preliminary designs approaches were
prototyped and

\subsubsection

@@

\section{Ground Segment}
\label{sec:ground-segment}

@@

\section{Testing}
\label{sec:testing}
